SNARL:\\Simulating Known Actors with Machine Learning

### 2017-05-15 tex/paper/paper.tex
Final paper and README.

We present SNARL, a suite of tools that is able to simulate a user's typing pattern with the goal of exposing the  vulnerabilities of new biometric based identification techniques. Researchers have known for many years that password based verifcation systems are insecure, in part due to security holes in applications, and in part due to users (eg password reuse). In an effort to bolster security, some applications leverage additional data; we are particularly interested in the merits and drawbacks of biometric data. Biometric identification schemes are in some aspects very secure, because a person can often be uniquely identified by some combination of biometric factors. On the other hand, if somehow this information were to be stolen, akin to a password being leaked, it is nearly impossible to change, since they are often either physical traits that exist from birth or habits developed over many years. Our paper demonstrates a system that is able to rapidly learn and simulate user password typing habits, and suggests that keypress timings should only be used as an inexpensive technique that does not significantly boost security.

@@

We present an exploration into the pitfalls of relying on input-delay-based biometric data as a second factor for authentication, using \texttt{punchTimeAuth} as our standard for comparison. We show that the model introduces a prohibively high number false negatives when exposed to real-world usage. Furthermore, we present serveral adversarial models which can successfully impersonate a user typing their password, trained on background typing data that need not include the password itself. In order to collect this data, we built two collection mediums: a Google Documents look-alike and an extension for Google Chrome. We collect which keys are pressed, the length of time each is held for, and the delay between each pair of keys. Finally, we present a novel neural model for attempting to distinguish users based on the information we collect, along with some conclusions and observations about this source of data as a second factor for security.

@@

The first is a Google Docs look-alike \cite{gdocs}, which imitates a well-known service in a phishing-style attempt to convince a complacent user to type on the page. We overlayed a text box on a webpage containing the CSS from a Google Doc we created, allowing the user to type as if this were the application they are so familiar with.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{admin}
  \caption{Data Debgging Dashboard}
\end{figure}

@@

The distinguisher model had a validation accuracy of 80\% to 85\%, which indicates that there is an appreciable difference between how distinct users type; however, it is clearly also not enought to uniquely identify a user. Again, this analysis suffers from low sample size, as we were only able to analyze four user's typing habits.

@@

We faced serval challenges in collecting data and implementing the models for this exploration. One of the largest barriers to rigorous results was the fact that the systems we attempted to break all require significant effort to train. Concretely, in order to evaluate an imitation of a user's typing of a specific password, that user must train the security system on how she types that password. This high barrier to the end user is a shortcoming of the string-specific model used in \texttt{punchTimeAuth}, and something we tried to address with our generalized neural distinguisher. However, as we discovered, simple typing timing and language context do not provide sufficient information to produce a generalized model of a user's biometric profile.

Another particularly interesting challenge occured in the client-side data collection library we authored. Detecting lengths of key presses in addition to delays meant we could not simply listen for \texttt{KeyPress} events, and put us at the mercy of the JavaScript event loop scheduler. We spent a nontrivial amount of time post-processing the stream of events we received to piece together the sequence of user key presses.

@@

We have shown that there is significant differentiation between user's typing patterns, as evidenced by the results of our distinguisher model. However, it is not clear to the authors that there exists sufficient entropy in this data to uniquely identify a user in a large system. Certainly, timing delays and key hold times do not render a model that should be used as a second factor alone. We would like to explore how the signal from such a model can be used a heuristic for other biometric factors, as well as whether or not such signals could be aggregated to provide a secure authenication system.

>>>>

We present SNARL, a suite of tools that is able to simulate a user's typing pattern with the goal of exposing the vulnerabilities of new biometric based identification techniques. Researchers have known for many years that password based verification systems are insecure, in part due to security holes in applications, and in part due to users (e.g. password reuse). In an effort to bolster security, some applications leverage additional data; we are particularly interested in the merits and drawbacks of biometric data. Biometric identification schemes are in some aspects very secure, because a person can often be uniquely identified by some combination of biometric factors. On the other hand, if somehow this information were to be stolen, akin to a password being leaked, it is nearly impossible to change, since they are often either physical traits that exist from birth or habits developed over many years. Our paper demonstrates a system that is able to rapidly learn and simulate user password typing habits, and suggests that keypress timings should only be used as an inexpensive technique that does not significantly boost security.

@@

We present an exploration into the pitfalls of relying on input-delay-based biometric data as a second factor for authentication, using \texttt{punchTimeAuth} as our standard for comparison. We show that the model introduces a prohibitively high number false negatives when exposed to real-world usage. Furthermore, we present several adversarial models which can successfully impersonate a user typing their password, trained on background typing data that need not include the password itself. In order to collect this data, we built two collection mediums: a Google Documents look-alike and an extension for Google Chrome. We collect which keys are pressed, the length of time each is held for, and the delay between each pair of keys. Finally, we present a novel neural model for attempting to distinguish users based on the information we collect, along with some conclusions and observations about this source of data as a second factor for security.

@@

The first is a Google Docs look-alike \cite{gdocs}, which imitates a well-known service in a phishing-style attempt to convince a complacent user to type on the page. We overlaid a text box on a webpage containing the CSS from a Google Doc we created, allowing the user to type as if this were the application they are so familiar with.

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{admin}
  \caption{Data Debugging Dashboard}
\end{figure}

@@

The distinguisher model had a validation accuracy of 80\% to 85\%, which indicates that there is an appreciable difference between how distinct users type; however, it is clearly also not enough to uniquely identify a user. Again, this analysis suffers from low sample size, as we were only able to analyze four user's typing habits.

@@

We faced several challenges in collecting data and implementing the models for this exploration. One of the largest barriers to rigorous results was the fact that the systems we attempted to break all require significant effort to train. Concretely, in order to evaluate an imitation of a user's typing of a specific password, that user must train the security system on how she types that password. This high barrier to the end user is a shortcoming of the string-specific model used in \texttt{punchTimeAuth}, and something we tried to address with our generalized neural distinguisher. However, as we discovered, simple typing timing and language context do not provide sufficient information to produce a generalized model of a user's biometric profile.

Another particularly interesting challenge occurred in the client-side data collection library we authored. Detecting lengths of key presses in addition to delays meant we could not simply listen for \texttt{KeyPress} events, and put us at the mercy of the JavaScript event loop scheduler. We spent a nontrivial amount of time post-processing the stream of events we received to piece together the sequence of user key presses.

@@

We have shown that there is significant differentiation between user's typing patterns, as evidenced by the results of our distinguisher model. However, it is not clear to the authors that there exists sufficient entropy in this data to uniquely identify a user in a large system. Certainly, timing delays and key hold times do not render a model that should be used as a second factor alone. We would like to explore how the signal from such a model can be used a heuristic for other biometric factors, as well as whether or not such signals could be aggregated to provide a secure authentication system.
